---
title: "Aides Drug"
author: "Marta Marriott"
date: "02/22/2024"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
triUmp <-read.csv(file="triUmpsal.csv",header=TRUE)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative aides drug (triUmpsal) is under test in clinical trials. Consider only the DrugUsers, does this drug seem effective?  Hint - use dplyr to obtain that subpopulation with the filter command.


```{r}
drugies <- filter(triUmp, DrugUser == "yes")
```


# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->



```{r}
barchartGC(~infected + treatment,data=drugies)
barchartGC(~infected + treatment,data=drugies, type="percent")

```


Nearly 80% of the patients who have aids, had been reported to have taken the placebo and not the genuine drug. That indicates the drug is effective in preventing patients from contracting aids.


### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->


```{r}
table1 <- xtabs(~infected + treatment, data=drugies)
rowPerc(table1)
colPerc(table1)
```


There were three times as many patients in  placebo group contracted aids, in comparison to the patients who received the genuine drug. 


## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->


```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

In this portion of the study three inferential tests were run, tow Chi-Squared tests and a Fisher's exact test. The Chi-Squared tests gave us a p value of 0.0000001242 which is far lower than a 0.05.and a Chi-Squared equal 27.955, or approximately 28, which is far greater than the critical degrees of freedom in this case 1. Given the particular results we can infer a difference in the data.


# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->


We can conclude from our graphical and numerical data, the patients in the placebo group contracted aids at a higher frequency than patients in the treatment group. That along with the statistical inference shows there is a statistical difference in the data. We can also conclude there is a negative correlation between the drug use the frequency of the disease. Due to the high number of participants, random variation did not have a significant role in the data. Therefore, we can infer the drug is effective in preventing aids.
